### 2023-04-05 get_descriptives_cohorts_nice.Rmd
Final cluster probabilities functions and how to use them

---
title: "functions and useage"
output: html_document
date: "2023-04-05"
---

```{r}
library(haven)
library(readr)
library(dplyr, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
library(purrr)
```

loading biomarker data 
```{r}
recoded_dat <- read_tsv("P:/FSE_MACSBIO/Smit, Femke/clustering project/data/recoded_dat.tsv", show_col_types = FALSE)
# [is.na(recoded_dis)] <- 0

strat_raw_dat <- recoded_dat %>% split( f = .$sex) %>% map(select, -c("sex"))
map(strat_raw_dat, head)

strat_dat <- recoded_dat %>%
  split(f = .$sex) %>%
  map(
    ~{
      .x %>%
        names %>%
        setdiff(c("eid", "sex", "age", "smoking", "bmi")) %>%
        map_dfc(
          function(feature){
            paste(feature, "~ age + smoking + bmi") %>%
              lm(data = .x) %>%
              resid %>%
              scale %>%
              data.frame %>%
              setNames(feature)
          }
        ) %>%
        mutate(eid = .x$eid, .before = 1)
    }
  )
map(strat_dat, head)
```

loading disease data
```{r}
recoded_dis <- read_tsv("P:/FSE_MACSBIO/Smit, Femke/clustering project/data/recoded_dis2.tsv", show_col_types = FALSE)  #any cancer is skin + other cancers
head(recoded_dis)

strat_dis <- merge(recoded_dat[,c("eid","sex")], recoded_dis, by = "eid") %>% split(f = .$sex) %>% map(select, -"sex")
map(strat_dis, head)
```

defining function cluster probabilities, cluster characteristics, cluster weighted disease prevalences
```{r}
getclusprob <- function(X, centers, covmats, weights){
    # Calculating probability density functions
    pdfs <- map2(centers, covmats, function(mu, covmat){ mvtnorm::dmvnorm(X, mu, covmat) })
    # Calculating likelihoods
    L <- map2(pdfs, weights, function(pd, w){ pd * w })
    # Joining in a matrix
    Lmat <- do.call(cbind, L)
    # Scaling by row to obtain probabilities
    probs <- Lmat / rowSums(Lmat)
    probs
}

getclusproball <- function(dataset, parameters){
  d <- select(dataset, -eid)
  cluspars <- parameters
  centers <- map(cluspars, pluck, "center")
  covmats <- map(cluspars, pluck, "cov")
  weights <- map(cluspars, pluck, "weight")
  getclusprob(d, centers, covmats, weights)
}

get_cluster_descriptives <- function(dataset, clus_probs){
  dataset <- data.frame(dataset)
  long_summary <- data.frame(cluster = integer(0), n_allocated = integer(0), perc_allocated = numeric(0), variable = character(0),  wtd_mean = numeric(0) , wtd_std = numeric(0), wtd_q0 = numeric(0), wtd_q25 = numeric(0), wtd_q50 = numeric(0), wtd_q75 = numeric(0), wtd_q100 = numeric(0))
  for (col in 1:ncol(clus_probs)){
    clus_name <- colnames(clus_probs)[col]
    probs_col <- clus_probs[,col]
    n_allocated <- sum(probs_col > 0.8)
    perc_allocated <- n_allocated/nrow(dataset)
    prob_sum <- sum(probs_col)
    for (var in 2:ncol(dataset)){
      var_name <- colnames(dataset)[var]
      var_col <- dataset[,var]
      wtd_mean <- sum(probs_col*var_col)/prob_sum
      wtd_std <- sum(probs_col*(var_col - wtd_mean))/prob_sum %>% sqrt()
      var_prob <- data.frame(var_col, probs_col)
      var_prob <- var_prob[order(var_prob$var_col),]
      var_prob$probs_cum <- cumsum(var_prob$probs_col)
      q0 <- var_prob$var_col[1]
      i25 <- which.min(abs(var_prob$probs_cum - 0.25*prob_sum))
      q25 <- var_prob$var_col[i25]
      i50 <- which.min(abs(var_prob$probs_cum - 0.5*prob_sum))
      q50 <- var_prob$var_col[i50]
      i75 <- which.min(abs(var_prob$probs_cum - 0.75*prob_sum))
      q75 <- var_prob$var_col[i75]
      q100 <- max(var_col)
      row <- nrow(long_summary) + 1
      long_summary[row,] <- c(clus_name, n_allocated, perc_allocated, var_name, wtd_mean, wtd_std, q0, q25, q50, q75, q100)
    }
  }
  long_summary
}

get_disease_prev <- function(dataset, clus_probs){
  #first column dataset needs to contain ids  
  dataset <- data.frame(dataset)
  long_summary <- data.frame(cluster = integer(0), n_allocated = integer(0), perc_allocated = numeric(0), variable = character(0),  wtd_prev = numeric(0))
  for (col in 1:ncol(clus_probs)){
    clus_name <- colnames(clus_probs)[col]
    probs_col <- clus_probs[,col]
    n_allocated <- sum(probs_col > 0.8)
    perc_allocated <- n_allocated/nrow(dataset)
    prob_sum <- sum(probs_col)
    for (var in 2:ncol(dataset)){
      var_name <- colnames(dataset)[var]
      var_col <- dataset[,var]
      wtd_prev <- sum(probs_col*var_col)/prob_sum*1000
      row <- nrow(long_summary) + 1
      long_summary[row,] <- c(clus_name, n_allocated, perc_allocated, var_name, wtd_prev)
    }
  }
  long_summary
}
```

getting cluster probabilities
```{r}
load("C:/Users/P70074150/Downloads/cluster_params_Maastricht.RData")
cluster_probs <- map2(strat_dat, cluster_params, getclusproball) 
map(cluster_probs, head)
```

getting cluster characteristics
```{r}
cluster_descriptives_resids <- map2(strat_dat, cluster_probs, get_cluster_descriptives)
map(cluster_descriptives_resids, head)

cluster_descriptives_raw <- map2(strat_raw_dat, cluster_probs, get_cluster_descriptives)
map(cluster_descriptives_raw, head)
```

getting disease prevalence
We can consider using this function also for smoking (and medicine use, etc, just all binary variables) as only the prevalence makes sense for these, there's no variance or uncertainty estimation relevant here.
```{r}
cluster_disease_prev <- map2(strat_dis, cluster_probs, get_disease_prev)
map(cluster_disease_prev, head)

```