### 2024-02-19 08_PrevDx.Rmd

---
title: An R Markdown document converted from "08_PrevDx.ipynb"
output: html_document
---

# Relation with prevalent diseases

```{r}
library(readr)
library(dplyr, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
library(purrr)
library(ggplot2)
library(Matrix,  warn.conflicts = FALSE)
library(lme4, warn.conflicts = FALSE)
library(lmerTest, warn.conflicts = FALSE)
```

```{r}
options(repr.plot.res = 300)
```

## Importing results

```{r}
result_tab <- tibble(
    cohort = c("ukb", "maastricht", "rotterdam", "ghs")
)
result_tab
```

```{r}
result_tab <- result_tab %>%
    mutate(
        data = map(
            cohort,
            ~tryCatch(rio::import(paste0("../data/", .x, "/result_file2.RData")),
                      error = \(e) NULL)
        )
    )
print(result_tab)
```

```{r}
result_crosssect <- result_tab %>%
    expand_grid(
        element_name = c(
            "MarkerDistrib",
            "BMIeffOnMarker",
            "CountCovars",
            "CountDXMeds",
            "CrossSectAssoc"
        )
    ) %>%
    transmute(
        cohort, element_name,
        element_value = map2(
            data, element_name,
            ~.x[[.y]]
        )
    )
print(result_crosssect)
```

```{r}
result_crosssect <- result_crosssect %>%
    group_by(element_name) %>%
    summarise(
        element_value = list(
            bind_rows(
                map2(
                    element_value, cohort, 
                    ~mutate(.x, cohort = .y, .before = 1)
                )
            )
        )
    )
print(result_crosssect)
```

## Color map of clusters

```{r}
load("../data/validclusmod.RData")
```

```{r}
print(validclusmod)
```

```{r}
cluscolmap <- validclusmod %>%
    transmute(
        clusmod = map(clusmod, select, Cluster = validclus_name, W = validclus_weights)
    ) %>%
    unnest(clusmod) %>%
    group_by(Cluster) %>%
    summarise(W2 = exp(sum(log((W))))) %>%
    arrange(-W2) %>%
    mutate(
        Cluster = factor(Cluster, levels = Cluster),
        ClusCol = scales::hue_pal()(n())
    )
cluscolmap
```

## Proportion of individuals in each cluster

```{r}
propclusall <- result_crosssect %>%
    filter(element_name == "MarkerDistrib") %>%
    unnest(element_value) %>%
    select(cohort, sex, Cluster, N) %>%
    unique %>%
    nest(D = -sex) %>%
    mutate(
        D = map(D, pivot_wider, names_from = Cluster, values_from = N)
    ) %>%
    unnest(D)
propclusall
```

```{r}
write_tsv(propclusall, "../data/propclusall.tsv")
```

```{r}
options(repr.plot.width = 5, repr.plot.height = 2)
propclusall %>%
    pivot_longer(-c(sex, cohort), names_to = "Cluster", values_to = "N", values_drop_na = TRUE) %>%
    group_by(sex, cohort) %>%
    inner_join(cluscolmap, by = "Cluster") %>%
    mutate(
        Cluster = factor(Cluster, levels = cluscolmap$Cluster),
        Nsum = sum(N),
        prop = N/Nsum,
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    ggplot(aes(interaction(cohort, Cluster), prop)) +
    geom_col(aes(fill = ClusCol)) +
    scale_fill_identity() +
    scale_x_discrete(guide = ggh4x::guide_axis_nested()) +
    scale_y_continuous(labels = scales::percent) +
    ggh4x::facet_nested(
        ~sex, 
        scales = "free",
        strip = ggh4x::strip_nested( 
            size = "variable"
        )
    ) +
    labs(x = NULL, y = NULL) + 
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 4, angle = 90),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 6),
        strip.text = element_text(size = 6),
        legend.position = "top"
    )
```

## Distribution of biomarkers

```{r}
bm_meansd <- result_crosssect %>%
    filter(element_name == "MarkerDistrib") %>%
    unnest(element_value) %>%
    filter(Type == "Numeric") %>%
    mutate(
        Summary1 = gsub("\\(|\\)", "", Summary1),
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    separate_wider_delim(Summary1, delim = " ", names = c("Mean", "SD")) %>%
    transmute(cohort, sex, Variable, Cluster, across(c(N, Mean, SD), as.numeric))
head(bm_meansd)
```

```{r}
bm_meansd %>%
    select(-N) %>%
    nest(D = -sex) %>%
    mutate(
        D = map(D, pivot_wider, names_from = Cluster, values_from = c(Mean, SD), names_vary = "slowest"),
        D = map2(D, sex, ~mutate(.x, sex = .y, .before = 1)),
        D = map(D, arrange, Variable)
    ) %>%
    pull(D)
```

```{r}
pooledmean <- bm_meansd %>%
    nest(D = -c(sex, Variable, Cluster)) %>%
    mutate(
        D = map(
            D,
            ~meta::metamean(n = .x$N, mean = .x$Mean, sd = .x$SD)
        ),
        D = map(
            D,
            ~tibble(
                N = sum(.x$n),
                Mean = .x$TE.fixed, 
                SD = sqrt(sum(.x$n)) * .x$seTE.fixed
            )
        )
    ) %>%
    unnest(D)
head(pooledmean)
```

```{r}
allmeansd <- pooledmean %>%
    mutate(cohort = "Pooled", .before = 1) %>%
    bind_rows(bm_meansd)
head(allmeansd)
```

```{r}
write_tsv(allmeansd, "../data/allmeansd.tsv")
```

```{r}
pooledmean %>%
    select(-N) %>%
    nest(D = -sex) %>%
    mutate(
        D = map(
            D, 
            pivot_wider, 
            names_from = Cluster, 
            values_from = c(Mean, SD), 
            names_vary = "slowest"
        ),
        D = map(D, arrange, Variable)
    ) %>%
    pull(D)
```

```{r}
bm_iqr <- result_crosssect %>%
    filter(element_name == "MarkerDistrib") %>%
    unnest(element_value) %>%
    filter(Type == "Numeric") %>%
    mutate(
        Summary2 = stringr::str_replace_all(
            Summary2,
            c(
                "\\(" = "- ",
                "\\)" = ""
            )
        ),
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    separate_wider_delim(
        Summary2, 
        delim = " - ", 
        names = paste(
            "calc",
            c(
                "Median",
                "minval",
                "p25",
                "p75",
                "maxval"
            ),
            sep = "_"
        )
    ) %>%
    transmute(
        cohort, sex, Variable, Cluster, N,
        across(starts_with("calc"), as.numeric)
    ) %>%
    rename_with(\(x) gsub("calc_", "", x), starts_with("calc"))
head(bm_iqr)
```

```{r}
write_tsv(bm_iqr, "../data/bm_iqr.tsv")
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 5)
bm_iqr %>%
    inner_join(cluscolmap, by = "Cluster") %>%
    mutate(
        Variable = toupper(Variable),
        Cluster = factor(Cluster, levels = rev(cluscolmap$Cluster))
    ) %>%
    ggplot(aes(y = interaction(cohort, Cluster))) +
    geom_boxplot(
        aes(
            xmin = minval,
            xlower = p25,
            xmiddle = Median,
            xupper = p75,
            xmax = maxval,
            fill = ClusCol
        ),
        stat = "identity",
        size = .5
    ) +
    scale_fill_identity() +
    scale_y_discrete(guide = ggh4x::guide_axis_nested()) +
    ggh4x::facet_nested(
        sex ~ Variable, 
        scales = "free",
        space = "free_y",
        strip = ggh4x::strip_nested(
            text_y = ggh4x::elem_list_text(angle = 0), 
            size = "variable"
        )
    ) +
    labs(x = NULL, y = NULL) + 
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 4),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 6),
        strip.text = element_text(size = 6)
    )
```

Prevalence of smoking:

```{r}
options(repr.plot.width = 7, repr.plot.height = 2.5)
result_crosssect %>%
    filter(element_name == "MarkerDistrib") %>%
    unnest(element_value) %>%
    filter(Variable == "smoking", Summary1 == 1) %>%
    mutate(
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        ),
        cohort = factor(cohort, levels = c(sort(unique(cohort), decreasing = TRUE))),
        prop = as.numeric(Summary2)/N
    ) %>%
    inner_join(cluscolmap, by = "Cluster") %>%
    mutate(
        Cluster = factor(Cluster, levels = cluscolmap$Cluster)
    ) %>%
    ggplot(aes(Cluster, prop)) +
    geom_col(aes(fill = ClusCol)) +
    scale_fill_identity() +
    scale_y_continuous(labels = scales::percent) +
    ggh4x::facet_nested(
        ~sex + cohort, 
        scales = "free",
        strip = ggh4x::strip_nested( 
            size = "variable"
        )
    ) +
    labs(x = NULL, y = NULL) + 
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 4, angle = 45, hjust = 0.5),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 6),
        strip.text = element_text(size = 6),
        legend.position = "top"
    )
```

## BMI effects by cluster

```{r}
bmieffdf <- result_crosssect %>%
    filter(element_name == "BMIeffOnMarker") %>%
    unnest(element_value) %>%
    filter(term == "bmi") %>%
    select(-c(element_name, term))
head(bmieffdf)
```

```{r}
pooledBMIEff <- bmieffdf %>%
    nest(D = -c(sex, Variable, Cluster)) %>%
    mutate(
        D = map(
            D,
            ~meta::metagen(TE = .x$estimate, seTE = .x$se)
        ),
        D = map(
            D,
            ~tibble(
                cohort = c("Pooled_FE", "Pooled_RE"),
                estimate = c(.x$TE.fixed, .x$TE.random), 
                se = c(.x$seTE.fixed, .x$seTE.random)
            )
        )
    ) %>%
    unnest(D)
head(pooledBMIEff)
```

```{r}
allbmieff <- pooledBMIEff %>%
    bind_rows(bmieffdf) %>%
    mutate(
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        ),
        pval = 2 * pnorm(-abs(estimate/se)),
        lwr = estimate - qnorm(.975)*se,
        upr = estimate + qnorm(.975)*se
    )
head(allbmieff)
```

```{r}
write_tsv(allbmieff, "../data/allbmieff.tsv")
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 8)
allbmieff %>%
    mutate(
        cohort = factor(
            cohort,
            levels = c("UKB", "MS", "RS", "GHS", "Pooled_FE", "Pooled_RE")
        ),
        Cluster = factor(Cluster, levels = cluscolmap$Cluster),
        Variable = toupper(Variable)
    ) %>%
    ggplot(aes(cohort, estimate)) +
    geom_hline(yintercept = 0, lty = "dashed") +
    geom_linerange(
        aes(
            ymin = lwr,
            ymax = upr
        )
    ) +
    geom_point(
        aes(shape = grepl("Pooled", cohort)),
        show.legend = FALSE
    ) +
    scale_shape_manual(
        values = c(20, 23)
    ) +
    ggh4x::facet_nested(
        Variable ~ sex + Cluster, 
        scales = "free",
        strip = ggh4x::strip_nested(
            text_y = ggh4x::elem_list_text(angle = 0), 
            size = "variable"
        )
    ) +
    labs(y = "Effect per BMI unit (95% CI)", x = NULL) + 
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 7, angle = 90),
        axis.text.y = element_text(size = 5),
        axis.title.x = element_text(size = 6),
        strip.text = element_text(size = 6)
    )
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 3)
allbmieff %>%
    mutate(
        Cluster = factor(Cluster, levels = rev(cluscolmap$Cluster)),
        Variable = toupper(Variable),
        Variable = case_match(
            Variable,
            "ALT" ~ paste(Variable, "(U/L)"),
            "CRP" ~ paste(Variable, "(mg/L)"),
            c("SBP", "DBP") ~ paste(Variable, "(mmHg)"),
            c("FG", "HDL", "LDL", "TG") ~ paste(Variable, "(mmol/L)"),
            "SCR" ~ paste(Variable, "(umol/L)"),
            "WHR" ~ paste(Variable, "(cm/cm)")
        )
    ) %>%
    filter(cohort == "Pooled_RE") %>%
    ggplot(aes(estimate, Cluster)) +
    geom_vline(xintercept = 0, lty = "dashed") +
    geom_rect(
        aes(
            xmin = ifelse(Cluster == "BC", lwr, NA), 
            xmax = ifelse(Cluster == "BC", upr, NA), 
            ymin = -Inf, ymax = Inf
        ), 
        fill = "pink", alpha = 0.8,
        na.rm = TRUE
    ) +
    geom_linerange(aes(xmin = lwr, xmax = upr)) +
    geom_point() +
    scale_x_continuous(n.breaks = 3, labels = function(x) ifelse(x == 0, "0", x)) +
    facet_grid(sex ~ Variable, scales = "free", space = "free_y") +
    labs(x = "Difference in biomarker per BMI unit increase", y = NULL) +
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 5),
        axis.title = element_text(size = 8)
    )
```

## ORs of diseases relative to concordant cluster

```{r}
percohortORs <- result_crosssect %>%
    filter(element_name == "CountCovars") %>%
    unnest(element_value) %>%
    filter(
        Covariate %in% c(
            "CHD", "Stroke", "T2D", "HT", "T1D", "LiverFailure", "RA"
        ),
        !(cohort %in% c("maastricht", "rotterdam") & 
          Covariate %in% c("LiverFailure", "RA", "T1D")),
        !(cohort == "ghs" & Covariate == "RA")
    ) %>%
    select(-c(element_name, Nclus)) %>%
    nest(D = -c(sex, cohort, Covariate)) %>%
    mutate(
        Covariate = factor(
            Covariate,
            levels = c(
                "CHD", "Stroke", "T2D", "HT", "T1D", "LiverFailure", "RA"
            )
        ),
        mod = map(
            D, 
            ~glm(
                cbind(round(Ncases), round(Nnoncases)) ~ Cluster, 
                data = .x, family = "binomial"
            )
        )
    ) %>%
    mutate(
        mod = map(mod, broom::tidy),
        mod = map(mod, mutate, term = gsub("Cluster", "", term)),
        D = map2(D, mod, left_join, by = c("Cluster" = "term")),
        mod = NULL
    ) %>%
    unnest(D) %>%
    arrange(sex, cohort, Covariate)
head(percohortORs)
```

```{r}
pooledORs <- result_crosssect %>%
    filter(element_name == "CountCovars") %>%
    unnest(element_value) %>%
    filter(
        Covariate %in% c(
            "CHD", "Stroke", "T2D", "HT", "T1D", "LiverFailure", "RA"
        ),
        !(cohort %in% c("maastricht", "rotterdam") & Covariate %in% c("LiverFailure", "RA")),
        !(cohort == "ghs" & Covariate == "RA")
    ) %>%
    select(-c(element_name, Nclus)) %>%
    nest(D = -c(sex, Covariate)) %>%
    mutate(
        Covariate = factor(
            Covariate,
            levels = c(
                "CHD", "Stroke", "T2D", "HT", "T1D", "LiverFailure", "RA"
            )
        ),
        Pooled_FE = map(
            D,
            ~if(length(unique(.x$cohort)) > 1){
                glm(
                    cbind(round(Ncases), round(Nnoncases)) ~ Cluster, 
                    data = .x, family = binomial
                )
            } else {
                NULL
            }                
        ),
        Pooled_RE = map(
            D,
            ~if(length(unique(.x$cohort)) > 1){
                glmer(
                    cbind(round(Ncases), round(Nnoncases)) ~ Cluster + (1|cohort), 
                    data = .x, family = binomial
                )
            } else {
                NULL
            }                
        )
    ) %>%
    pivot_longer(c(Pooled_FE, Pooled_RE), names_to = "cohort", values_to = "mod") %>%
    filter(!map_lgl(mod, is.null)) %>%
    mutate(
        mod = map2(cohort, mod, ~if(.x == "Pooled_FE"){ broom::tidy(.y) } else { broom.mixed::tidy(.y) }),
        mod = map(mod, mutate, term = gsub("Cluster", "", term)),
        D = map(D, group_by, Cluster),
        D = map(D, summarise, across(c(Ncases, Nnoncases), sum)),
        D = map2(D, mod, left_join, by = c("Cluster" = "term")),
        mod = NULL
    ) %>%
    unnest(D) %>%
    select(-c(effect, group)) %>%
    arrange(sex, Covariate)
head(pooledORs)
```

```{r}
allORs <- pooledORs %>%
    bind_rows(percohortORs) %>%
    transmute(
        sex, Covariate, 
        cohort = factor(
            case_match(
                cohort,
                "ukb" ~ "UKB",
                "maastricht" ~ "MS",
                "rotterdam" ~ "RS",
                "ghs" ~ "GHS",
                .default = cohort
            ),
            levels = c("UKB", "MS", "RS", "GHS", "Pooled_FE", "Pooled_RE")
        ),      
        Cluster = gsub("prob", "", Cluster), 
        across(everything()),
        OR = exp(estimate),
        LOWERCI = exp(estimate - qnorm(1 - .05/2)*std.error),
        UPPERCI = exp(estimate + qnorm(1 - .05/2)*std.error)
    ) %>%
    arrange(sex, Covariate, Cluster, cohort)
head(allORs)
nrow(allORs)
```

```{r}
write_tsv(allORs, "../data/allORs.tsv")
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 3)
allORs %>%
    drop_na %>%
    filter(cohort %in% c("Pooled_FE", "Pooled_RE") | Covariate == "RA") %>%
    mutate(
        cohort = gsub("Pooled_", "", cohort),
        Cluster = factor(Cluster, levels = rev(cluscolmap$Cluster))
    ) %>%
    ggplot(aes(OR, Cluster)) +
    geom_vline(xintercept = 1, lty = "dashed") +
    geom_linerange(
        aes(xmin = LOWERCI, xmax = UPPERCI, group = cohort),
        position = position_dodge(width = .5)
    ) +
    geom_point(
        aes(group = cohort, color = cohort),
        position = position_dodge(width = .5)
    ) +
    facet_grid(sex ~ Covariate, scales = "free", space = "free_y") +
    labs(x = "OR relative to BC", y = NULL, color = "Model") +
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 5)
    )
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 3)
allORs %>%
    drop_na %>%
    filter(cohort == "Pooled_RE" | Covariate == "RA") %>%
    mutate(
        bigeff = OR > 10, xint = ifelse(bigeff, NA, 1),
        Cluster = factor(Cluster, levels = rev(cluscolmap$Cluster))
    ) %>%
    ggplot(aes(OR, Cluster)) +
    geom_vline(aes(xintercept = xint), lty = "dashed", na.rm = TRUE) +
    geom_linerange(
        aes(xmin = LOWERCI, xmax = UPPERCI)
    ) +
    geom_point() +
    ggh4x::facet_nested(
        sex ~ Covariate + bigeff, scales = "free", space = "free_y",
        strip = ggh4x::strip_nested(
            size = "variable",
            background_x = list(element_rect(), element_rect(fill = NA, colour = NA)),
            text_x = list(element_text(), element_text(size = 0, margin = margin())),
            by_layer_x = TRUE
        )
    ) +
    ggh4x::force_panelsizes(cols = c(1, 1, 1, .5, 1, 1, .5, 1, 1)) +
    ggh4x::scale_x_facet(!bigeff, n.breaks = 4, limits = c(0, NA)) +
    ggh4x::scale_x_facet(bigeff, n.breaks = 4) +
    labs(x = "OR relative to BC", y = NULL, color = "Model") +
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 5)
    )
```

## ORs of medications relative to concordant cluster

```{r}
percohortORmeds <- result_crosssect %>%
    filter(element_name == "CountCovars") %>%
    unnest(element_value) %>%
    filter(
        Covariate %in% c(
            "Insulin", "AntiHT", "LipidLower"
        ),
        !(cohort %in% c("maastricht", "rotterdam") & Covariate == "Insulin")
    ) %>%
    select(-c(element_name, Nclus)) %>%
    nest(D = -c(sex, cohort, Covariate)) %>%
    mutate(
        Covariate = factor(
            Covariate,
            levels = c(
                "Insulin", "AntiHT", "LipidLower"
            )
        ),
        mod = map(
            D, 
            ~glm(
                cbind(round(Ncases), round(Nnoncases)) ~ Cluster, 
                data = .x, family = "binomial"
            )
        )
    ) %>%
    mutate(
        mod = map(mod, broom::tidy),
        mod = map(mod, mutate, term = gsub("Cluster", "", term)),
        D = map2(D, mod, left_join, by = c("Cluster" = "term")),
        mod = NULL
    ) %>%
    unnest(D) %>%
    arrange(sex, cohort, Covariate)
head(percohortORmeds)
```

```{r}
pooledORmeds <- result_crosssect %>%
    filter(element_name == "CountCovars") %>%
    unnest(element_value) %>%
    filter(
        Covariate %in% c(
            "Insulin", "AntiHT", "LipidLower"
        ),
        !(cohort %in% c("maastricht", "rotterdam") & Covariate == "Insulin")
    ) %>%
    select(-c(element_name, Nclus)) %>%
    nest(D = -c(sex, Covariate)) %>%
    mutate(
        Covariate = factor(
            Covariate,
            levels = c(
                "Insulin", "AntiHT", "LipidLower"
            )
        ),
        Pooled_FE = map(
            D,
            ~if(length(unique(.x$cohort)) > 1){
                glm(
                    cbind(round(Ncases), round(Nnoncases)) ~ Cluster, 
                    data = .x, family = binomial
                )
            } else {
                NULL
            }                
        ),
        Pooled_RE = map(
            D,
            ~if(length(unique(.x$cohort)) > 1){
                glmer(
                    cbind(round(Ncases), round(Nnoncases)) ~ Cluster + (1|cohort), 
                    data = .x, family = binomial
                )
            } else {
                NULL
            }                
        )
    ) %>%
    pivot_longer(c(Pooled_FE, Pooled_RE), names_to = "cohort", values_to = "mod") %>%
    filter(!map_lgl(mod, is.null)) %>%
    mutate(
        mod = map2(cohort, mod, ~if(.x == "Pooled_FE"){ broom::tidy(.y) } else { broom.mixed::tidy(.y) }),
        mod = map(mod, mutate, term = gsub("Cluster", "", term)),
        D = map(D, group_by, Cluster),
        D = map(D, summarise, across(c(Ncases, Nnoncases), sum)),
        D = map2(D, mod, left_join, by = c("Cluster" = "term")),
        mod = NULL
    ) %>%
    unnest(D) %>%
    select(-c(effect, group)) %>%
    arrange(sex, Covariate)
head(pooledORmeds)
```

```{r}
allORmeds <- pooledORmeds %>%
    bind_rows(percohortORmeds) %>%
    transmute(
        sex, Covariate, 
        cohort = factor(
            case_match(
                cohort,
                "ukb" ~ "UKB",
                "maastricht" ~ "MS",
                "rotterdam" ~ "RS",
                "ghs" ~ "GHS",
                .default = cohort
            ),
            levels = c("UKB", "MSS", "RS", "GHS", "Pooled_FE", "Pooled_RE")
        ),       
        Cluster = gsub("prob", "", Cluster), 
        across(everything()),
        OR = exp(estimate),
        LOWERCI = exp(estimate - qnorm(1 - .05/2)*std.error),
        UPPERCI = exp(estimate + qnorm(1 - .05/2)*std.error)
    ) %>%
    arrange(sex, Covariate, Cluster, cohort)
head(allORmeds)
nrow(allORmeds)
```

```{r}
write_tsv(allORmeds, "../data/allORmeds.tsv")
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 3)
allORmeds %>%
    drop_na %>%
    filter(cohort %in% c("Pooled_FE", "Pooled_RE")) %>%
    mutate(
        cohort = gsub("Pooled_", "", cohort),
        Cluster = factor(Cluster, levels = rev(cluscolmap$Cluster))
    ) %>%
    ggplot(aes(OR, Cluster)) +
    geom_vline(xintercept = 1, lty = "dashed") +
    geom_linerange(
        aes(xmin = LOWERCI, xmax = UPPERCI, group = cohort),
        position = position_dodge(width = .5)
    ) +
    geom_point(
        aes(group = cohort, color = cohort),
        position = position_dodge(width = .5)
    ) +
    facet_grid(sex ~ Covariate, scales = "free", space = "free_y") +
    labs(x = "OR relative to BC", y = NULL, color = "Model") +
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 5)
    )
```

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
allORmeds %>%
    drop_na %>%
    filter(cohort == "Pooled_RE") %>%
    mutate(
        bigeff = OR > 10, xint = ifelse(bigeff, NA, 1),
        Cluster = factor(Cluster, levels = rev(cluscolmap$Cluster))
    ) %>%
    ggplot(aes(OR, Cluster)) +
    geom_vline(aes(xintercept = xint), lty = "dashed", na.rm = TRUE) +
    geom_linerange(
        aes(xmin = LOWERCI, xmax = UPPERCI)
    ) +
    geom_point() +
    ggh4x::facet_nested(
        sex ~ Covariate + bigeff, scales = "free", space = "free_y",
        strip = ggh4x::strip_nested(
            size = "variable",
            background_x = list(element_rect(), element_rect(fill = NA, colour = NA)),
            text_x = list(element_text(), element_text(size = 0, margin = margin())),
            by_layer_x = TRUE
        )
    ) +
    ggh4x::force_panelsizes(cols = c(1, .6, 1, 1)) +
    ggh4x::scale_x_facet(!bigeff, limits = c(0, NA)) +
    ggh4x::scale_x_facet(bigeff, breaks = c(30, 40, 50)) +
    labs(x = "OR relative to BC", y = NULL, color = "Model") +
    theme_bw() +
    theme(
        axis.text.x = element_text(size = 5)
    )
```

## Proportion of people with diseases that are medicated

```{r}
dxmeddf <- result_crosssect %>%
    filter(element_name %in% c("CountCovars", "CountDXMeds")) %>%
    pivot_wider(names_from = element_name, values_from = element_value) %>%
    mutate(
        CountCovars = map(
            CountCovars, filter, 
            Covariate %in% c("CHD", "Stroke", "T2D")
        ),
        CountCovars = map(
            CountCovars, 
            pivot_longer, c(Ncases, Nnoncases), 
            names_to = "DxStatus", values_to = "DxStatusN"
        ),
        CountDXMeds = map(
            CountDXMeds, filter, 
            Dx %in% c("CHD", "T2D"),
            Med %in% c("NoMed", "AntiHT", "LipidLower")
        ),
        CountDXMeds = map(CountDXMeds, pivot_longer, c(Ncases, Nnoncases), 
                          names_to = "DxStatus", values_to = "DxMedInN"),
        CountDXMeds = map(CountDXMeds, rename, Covariate = Dx),
        joindat = map2(CountCovars, CountDXMeds, inner_join)
    ) %>%
    select(joindat) %>%
    unnest(joindat) %>%
    mutate(
        DxMedOutN = DxStatusN - DxMedInN,
        DxStatus = ifelse(DxStatus == "Ncases", 1, 0),
        across(c(Nclus, DxStatusN), \(x){ NULL })
    )
head(dxmeddf)
```

```{r}
dxmedmods <- dxmeddf %>%
    nest(D = -c(cohort, sex, Med, Covariate)) %>%
    mutate(
        model = map(D, ~glm(cbind(round(DxMedInN), round(DxMedOutN)) ~ Cluster * DxStatus, data = .x, family = binomial)),
        model = map(model, broom::tidy),
        D = NULL
    ) %>%
    unnest(model) %>%
    filter(term != "(Intercept)") %>%
    mutate(
        OR = exp(estimate),
        LOWERCI = exp(estimate - qnorm(1 - .05/2)*std.error),
        UPPERCI = exp(estimate + qnorm(1 - .05/2)*std.error)
    )
head(dxmedmods)
nrow(dxmedmods)
```

```{r}
pooldxmedmods <- dxmeddf %>%
    nest(D = -c(sex, Med, Covariate)) %>%
    mutate(
        Pooled_FE = map(
            D, 
            ~glm(
                cbind(round(DxMedInN), round(DxMedOutN)) ~ Cluster * DxStatus, 
                data = .x, family = binomial
            )
        ),
        Pooled_RE = map(
            D, 
            ~glmer(
                cbind(round(DxMedInN), round(DxMedOutN)) ~ Cluster * DxStatus + (1|cohort), 
                data = .x, family = binomial
            )
        ),
        D = NULL
    ) %>%
    pivot_longer(c(Pooled_FE, Pooled_RE), names_to = "cohort", values_to = "model") %>%
    mutate(
        model = map2(cohort, model, ~if(.x == "Pooled_FE"){ broom::tidy(.y) } else { broom.mixed::tidy(.y) })
    ) %>%
    unnest(model) %>%
    filter(!grepl("(Intercept)", term)) %>%
    select(-c(effect, group)) %>%
    mutate(
        OR = exp(estimate),
        LOWERCI = exp(estimate - qnorm(1 - .05/2)*std.error),
        UPPERCI = exp(estimate + qnorm(1 - .05/2)*std.error),
    )
head(pooldxmedmods)
```

```{r}
alldxmedmods <- pooldxmedmods %>%
    mutate(cohort = "Pooled") %>%
    bind_rows(dxmedmods) %>%
    transmute(
        sex, Covariate, 
        cohort = factor(
            case_match(
                cohort,
                "ukb" ~ "UKB",
                "maastricht" ~ "MS",
                "rotterdam" ~ "RS",
                "ghs" ~ "GHS",
                .default = cohort
            ),
            levels = c("UKB", "MS", "RS", "GHS", "Pooled_FE", "Pooled_RE")
        ),      
        term = gsub("prob", "", term), 
        across(everything())
    ) %>%
    arrange(sex, Covariate, term, Med, cohort)
head(alldxmedmods)
nrow(alldxmedmods)
```

```{r}
write_tsv(alldxmedmods, "../data/alldxmedmods.tsv")
```