### 2024-02-19 15_GRSAssoc.Rmd

---
title: An R Markdown document converted from "15_GRSAssoc.ipynb"
output: html_document
---

```{r}
library(readr)
library(dplyr, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
library(purrr)
library(ggplot2)
library(Matrix, warn.conflicts = FALSE)
library(uwot)
options(repr.plot.res = 300)
```

```{r}
sample_file <- "~/projects/DVA/Data/UKBiobank/ukb_57232/genetic_files/ukb57232_imp_chr1_v3_s487266.sample"
sample_file <- paste("tail -n+3", sample_file, "| cut -d' ' -f1") %>%
    pipe %>%
    read_tsv(col_types = "n", col_names = "eid")
head(sample_file)
nrow(sample_file)
```

```{r}
grs_tab <- map_dfc(list.files("../data/PRScoresUKB", full.names = TRUE),
                   read_tsv, show_col_types = FALSE) %>%
    mutate(eid = sample_file$eid, .before = 1)
head(grs_tab)
```

```{r}
grs_tab <- grs_tab %>%
    mutate(across(-eid, ~as.vector(scale(.x))))
head(grs_tab)
```

```{r}
genpcdat <- {
    tab_name <- "/ludc/Home/daniel_c/projects/DVA/Data/UKBiobank/ukb_57232/data_files/41806.tab"
    tab_vars <- tibble(
        colid = c("f.eid", paste0("f.22009.0.", 1:10)),
        colnam = c("eid", paste0("gpc", 1:10))
    )
    tab_vars <- paste("head -1", tab_name, "| tr '\t' '\n'") %>%
        pipe %>%
        read_table(col_names = "colid", col_types = "c") %>%
        mutate(colpos = row_number()) %>%
        inner_join(tab_vars, by = "colid")
    column_cmd <- paste("cut -f", paste(tab_vars$colpos, collapse = ","), tab_name)
    read_tsv(
        pipe(column_cmd),
        skip = 1, col_names = tab_vars$colnam,
        col_types = cols(.default = "n")
    )
}
head(genpcdat)
```

```{r}
load("../data/validclusmod.RData")
print(validclusmod)
```

```{r}
load("../data/ukb/strat_dat.RData")
print(strat_dat)
```

```{r}
validclusmod <- validclusmod %>%
    mutate(
        bmtab = map(sex, ~strat_dat[[.x]]),
        umapmod = map(
            sex, 
            ~load_uwot(paste0("../data/ukb/umapmod_", .x))
        ),
        residtab = map(
            umapmod,
            ~map(
                1:.x$nn_index$ann$getNItems(),
                function(i){
                    .x$nn_index$ann$getItemsVector(i-1)
                }
            )
        ),
        residtab = map(residtab, do.call, what = rbind),
        probtab = map2(
            residtab, clusmod,
            function(RESIDDF, CLUSMOD){
                pdfs <- mapply(
                    function(mu, covmat){ mvtnorm::dmvnorm(RESIDDF, mu, covmat) },
                    CLUSMOD$validclus_centers, 
                    CLUSMOD$validclus_covmats
                )
                L <- pdfs %*% diag(CLUSMOD$validclus_weights)
                colnames(L) <- paste0("prob", CLUSMOD$validclus_name)
                tibble(data.frame(L / rowSums(L)))
            }
        ),
        prsprobtab = map(
            probtab,
            mutate,
            across(everything(), ~ifelse(.x < 1e-15, 1e-15, .x)),
            across(everything(), ~ifelse(.x > 1 - 1e-15, 1 - 1e-15, .x)),
            across(-probBC, ~log(.x/probBC)),
            probBC = NULL
        ),
        prsprobtab = map2(
            bmtab, prsprobtab,
            ~bind_cols(select(.x, eid), .y)
        ),
        prsprobtab = map(
            prsprobtab,
            ~inner_join(.x, grs_tab, by = "eid")
        ),
        prsprobtab = map(
            prsprobtab,
            ~inner_join(.x, genpcdat, by = "eid")
        ),
        prsprobtab = map(
            prsprobtab,
            pivot_longer,
            starts_with("prob"),
            names_to = "Cluster",
            values_to = "logratio"
        ),
        prsprobtab = map(
            prsprobtab,
            mutate,
            Cluster = gsub("prob", "", Cluster)
        ),
        prsprobtab = map(
            prsprobtab,
            pivot_longer,
            -c(eid, Cluster, logratio, starts_with("gpc")),
            names_to = "GRSName",
            values_to = "GRSVal"
        )
    )
print(validclusmod)
```

```{r}
validclusmod$prsprobtab[[1]] %>% head
```

```{r}
clus_prs <- validclusmod %>%
    transmute(sex, prsprobtab = map(prsprobtab, nest, Data = -c(Cluster, GRSName))) %>%
    unnest(prsprobtab)
print(clus_prs)
```

```{r}
head(clus_prs$Data[[1]])
```

```{r}
clus_prs <- clus_prs %>%
    mutate(Mod = map(Data, ~lm(logratio ~ GRSVal + gpc1 + gpc2 + gpc3 + gpc4 + gpc5 + gpc6 + gpc7 + gpc8 + gpc9 + gpc10, data = .x)))
print(clus_prs)
```

```{r}
clus_prs_est <- clus_prs %>%
    transmute(
        sex, Cluster, Trait = GRSName,
        ModRes = map(Mod, broom::tidy, conf.int = TRUE)
    ) %>%
    unnest(ModRes) %>%
    filter(term == "GRSVal")
head(clus_prs_est)
```

```{r}
options(repr.plot.height = 3.5, repr.plot.width = 12)
clus_prs_est %>%
    mutate(
        across(c(estimate, conf.low, conf.high), exp),
        padj = p.adjust(p.value, "fdr")
    ) %>%
    ggplot(aes(estimate, Cluster)) +
    geom_vline(xintercept = 1, lty = "dashed") +
    geom_linerange(aes(xmin = conf.low, xmax = conf.high)) +
    geom_point(aes(color = padj < 0.05)) +
    facet_grid(sex ~ Trait, scales = "free", space = "free_y") +
    labs(
        x = "Odds ratio of having a discordant vs concordant profile\nper SD unit of PRS", 
        y = NULL,
        color = "p < 0.05 after 5% FDR correction"
    ) +
    theme_bw() +
    theme(
        legend.position = "top",
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 6),
        axis.text.x = element_text(size = 5.5),
        axis.title.x = element_text(size = 8)
    )
```