### 2024-07-25 scripts/16_PrevDxAdj.Rmd
Demo and other updates

---
title: An R Markdown document converted from "20_PrevDxAdj.ipynb"
output: html_document
---

# Associations with prevalent conditions adjusted for other prevalent conditions

## Libraries

```{r}
library(readr)
library(dplyr, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
library(purrr)
library(ggplot2)
```

```{r}
options(repr.plot.res = 300)
```

## Pre-defined functions

```{r}
source("06_cross_sectional_FX.R")
```

## Stratified data

```{r}
load("../data/ukb/strat_dat.RData")
```

```{r}
map(strat_dat, head)
```

## Validated clusters

```{r}
load("../data/validclusmod.RData")
print(validclusmod)
```

## Pre-existing conditions and medications

```{r}
covar_dat <- read_tsv("../data/covar_dat.tsv", show_col_types = FALSE)
head(covar_dat)
```

## Calculation of cluster probabilities

```{r}
clusterdfs <- clusterprobcalc(ClusModDf = validclusmod, StratDat = strat_dat)
```

```{r}
print(clusterdfs)
```

## Adding covariates

```{r}
clusterdfs <- addcovardat(X = clusterdfs, CovarDat = covar_dat)
```

```{r}
print(clusterdfs)
```

## Counting cases for each combination

```{r}
countdf <- clusterdfs %>%
    expand_grid(
        Adj = c("Un", "Dx", "Tx")
    ) %>%
    mutate(
        data = map(
            data, pivot_longer, starts_with("prob"),
            names_to = "Cluster",
            values_to = "ClusProb"
        ),
        data = map2(
            Adj, data,
            ~switch(
                .x,
                Un = group_by(.y, CHD, Cluster),
                Dx = group_by(.y, CHD, Cluster, HT, T2D, LipidLower),
                Tx = group_by(.y, CHD, Cluster, AntiHT, AntiDM, Insulin, LipidLower)
            )
        ),
        data = map(
            data,
            summarise,
            Nind = sum(ClusProb),
            .groups = "drop"
        ),
        data = map(
            data,
            pivot_wider,
            names_from = CHD, 
            values_from = Nind, 
            names_prefix = "CHD_"
        )
    )
print(countdf)
```

## Binomial regressions

```{r}
binreg <- countdf %>%
    mutate(
        data = map(
            data,
            ~glm(cbind(round(CHD_1), round(CHD_0)) ~ ., data = .x, family = binomial)
        ),
        data = map(
            data,
            broom::tidy
        )
    ) %>%
    unnest(data) %>%
    filter(grepl("Cluster", term)) %>%
    transmute(
        sex, Adj,
        Cluster = gsub("Clusterprob", "", term),
        OR = exp(estimate),
        LOWERCI = exp(estimate - qnorm(1 - .05/2)*std.error),
        UPPERCI = exp(estimate + qnorm(1 - .05/2)*std.error)
    )
head(binreg)
```

## Cluster map

```{r}
cluscolmap <- validclusmod %>%
    transmute(
        clusmod = map(clusmod, select, Cluster = validclus_name, W = validclus_weights)
    ) %>%
    unnest(clusmod) %>%
    group_by(Cluster) %>%
    summarise(W2 = exp(sum(log((W))))) %>%
    arrange(-W2) %>%
    mutate(
        Cluster = factor(Cluster, levels = Cluster),
        ClusCol = scales::hue_pal()(n())
    )
cluscolmap
```

## Plotting results

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
binreg %>%
    mutate(
        Cluster = factor(Cluster, levels = rev(cluscolmap$Cluster)),
        Adj = factor(
            Adj,
            levels = c("Un", "Dx", "Tx"),
            labels = c("None", "With\ncomorbidities", "With\ntreatment")
        ),
    ) %>%
    ggplot(aes(OR, Cluster)) +
    geom_vline(xintercept = 1, lty = "dashed", na.rm = TRUE) +
    geom_linerange(
        aes(xmin = LOWERCI, xmax = UPPERCI, group = Adj),
        position = position_dodge(width = .5), linewidth = .5
    ) +
    geom_point(
        aes(color = Adj, group = Adj), size = .75,
        position = position_dodge(width = .5)
    ) +
    scale_color_manual(
        values = c("black", "red", "green")
    ) +
    facet_grid(
        ~sex, scales = "free", space = "free_y"
    ) +
    labs(x = "CAD OR relative to BC", y = NULL, color = "Adjustment") +
    theme_bw() +
    theme(
        legend.position = "top"
    )
```