### 2025-11-12 scripts/02_ValClus.Rmd

---
title: An R Markdown document converted from "02_ValClus.ipynb"
output: html_document
---

# Validation results

```{r}
library(readr)
library(dplyr, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
library(purrr)
library(ggplot2)
```

```{r}
options(repr.plot.res = 300)
```

## Importing results

```{r}
result_tab <- tibble(
    cohort = c("ukb", "maastricht", "rotterdam", "ghs")
)
result_tab
```

```{r}
result_tab <- result_tab %>%
    mutate(
        data = map(
            cohort,
            ~rio::import(paste0("../data/", .x, "/result_file.RData"))
        )
    )
print(result_tab)
```

## General descriptives

```{r}
gendesc <- result_tab %>%
    transmute(
        cohort,
        data = map(data, "General_descriptives")
    ) %>%
    unnest(data)
head(gendesc)
```

### Table summarising biomarkers from all cohorts

```{r}
numgendesc <- gendesc %>%
    filter(Type == "Numeric") %>%
    mutate(across(c(Summary1, Summary2), ~stringr::str_replace_all(.x, c(" \\(" = " - ", "\\)$" = "")))) %>%
    separate_wider_delim(cols = Summary1, delim = " - ", names = c("Mean", "SD")) %>%
    separate_wider_delim(cols = Summary2, delim = " - ", names = c("Median", paste0("p", c(2.5, 25, 75, 97.5)))) %>%
    mutate(across(c(Mean, SD, Median, p2.5, p25, p75, p97.5), as.numeric))
head(numgendesc)
```

```{r}
write_tsv(numgendesc, "../data/numgendesc.tsv")
```

```{r}
numgendesc %>%
    select(cohort, sex, N) %>%
    unique
```

```{r}
gendesc %>%
    filter(Type == "Categorical")
```

```{r}
percsmoke <- gendesc %>%
    filter(Type == "Categorical") %>%
    transmute(
        cohort, sex,
        currentsmoke = ifelse(Summary1 == 0, "No", "Yes"),
        Summary2 = stringr::str_replace_all(Summary2, c(" \\(" = " - ", "%\\)$" = ""))
    ) %>%
    separate_wider_delim(
        cols = Summary2, delim = " - ", 
        names = c("Ncat", "Perc")
    ) %>%
    mutate(across(c(Ncat, Perc), as.numeric))
head(percsmoke)
```

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
percsmoke %>%
    filter(currentsmoke == "Yes") %>%
    mutate(
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "TMS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    ggplot(aes(cohort, Perc)) +
    geom_col(aes(fill = cohort), show.legend = FALSE) +
    scale_fill_viridis_d(option = "A", begin = .25, end = .75, direction = -1) +
    facet_wrap(
        ~sex
    ) +
    labs(x = NULL, y = "%") + 
    theme_bw()
```

### Boxplot summarising biomarkers from all cohorts

```{r}
options(repr.plot.width = 10, repr.plot.height = 5)
numgendesc %>%
    mutate(
        Variable = toupper(Variable),
        trait = Variable,
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "TMS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    group_by(Variable) %>%
    nest %>%
    mutate(
        data = map(
            data,
            ~.x %>%
                ggplot(aes(cohort)) +
                geom_boxplot(aes(fill = cohort, ymin = p2.5, lower = p25, middle = Median, upper = p75, ymax = p97.5),
                             stat = "identity") +
                scale_fill_viridis_d(option = "A", begin = .25, end = .75, direction = -1) +
                ggh4x::facet_nested_wrap(~trait + sex, nrow = 1) +
                theme_bw() +
                theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())
        )
    ) %>%
    pull(data) %>%
    patchwork::wrap_plots(nrow = 2, guides = "collect") &
    theme(legend.position = "top", legend.title = element_blank())
```

## Regression coefficients

```{r}
regcoefs <- result_tab %>%
    transmute(
        cohort,
        data = map(data, "BMI_coefficients")
    ) %>%
    unnest(data)
head(regcoefs)
```

```{r}
write_tsv(regcoefs, "../data/regcoefs.tsv")
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 5)
regcoefs %>%
    filter(term == "bmi") %>%
    mutate(
        Biomarker = toupper(Biomarker),
        trait = Biomarker,
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "TMS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    nest(data = -Biomarker) %>%
    mutate(
        data = map(
            data,
            ~.x %>%
                ggplot(aes(cohort, Estimate)) +
                geom_hline(yintercept = 0, lty = "dashed") +
                geom_linerange(aes(ymin = lowerCI, ymax = upperCI)) +
                geom_point(aes(color = cohort)) +
                scale_color_viridis_d(option = "A", begin = .25, end = .75, direction = -1) +
                ggh4x::facet_nested_wrap(~trait + sex, nrow = 1) +
                theme_bw() +
                theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title = element_blank())
        )
    ) %>%
    pull(data) %>%
    patchwork::wrap_plots(nrow = 2, guides = "collect") &
    theme(legend.position = "top")
```

## UMAP projections

```{r}
umap_projs <- result_tab %>%
    transmute(
        cohort,
        data = map(data, "Cluster_results"),
        sex = map(data, names)
    ) %>%
    unnest(c(sex, data)) %>%
    mutate(data = map(data, pluck, "umap_model", "embedding"),
           data = map(data, data.frame)) %>%
    unnest(data)
print(umap_projs)
```

Discovery and validation cohorts:

```{r}
options(repr.plot.width = 10, repr.plot.height = 6)
umap_projs %>%
    mutate(
        cohort = recode_factor(
            factor(cohort), 
            "ukb" = "UKB",
            "rotterdam" = "RS",
            "maastricht" = "TMS",
            "ghs" = "GHS"
        )
    ) %>%
    ggplot(aes(X1, X2)) +
    geom_point(aes(color = cohort), shape = ".", show.legend = FALSE) +
    scale_color_viridis_d(option = "A", begin = .25, end = .75, direction = 1) +
    ggh4x::facet_nested_wrap(~cohort + sex, scales = "free", nrow = 2) +
    theme_bw() +
    theme(
        axis.text = element_blank(),
        axis.ticks = element_blank()
    ) +
    labs(x = "UMAP1", y = "UMAP2")
```

## Clusters

```{r}
options(repr.plot.width = 12, repr.plot.height = 15)
result_tab %>%
    transmute(
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "TMS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        ),
        data = map(data, "Cluster_results"),
        sex = map(data, names)
    ) %>%
    unnest(c(sex, data)) %>%
    mutate(
        data = map(data, "clusters"),
        cluster = map(data, names)
    ) %>%
    unnest(cluster) %>%
    mutate(
        data = map2(data, cluster, ~.x[[.y]]),
        center = map(data, "center"),
        trait = map(center, names)
    ) %>%
    select(-data) %>%
    unnest(c(center, trait)) %>%
    mutate(trait = toupper(trait)) %>%
    ggplot(aes(center, trait)) +
    geom_vline(xintercept = 0, lty = "dashed") +
    geom_segment(aes(xend = 0, yend = trait)) +
    geom_point() +
    ggh4x::facet_nested(sex + cohort ~ cluster) +
    labs(x = "Mean residual value (SD units)", y = NULL) +
    theme_bw()
```

## Individuals allocated to their closest cluster

```{r}
closest_cluster <- result_tab %>%
    transmute(
        cohort,
        data = map(data, "Cluster_results"),
        sex = map(data, names)
    ) %>%
    unnest(c(sex, data)) %>%
    mutate(
        embed = map(data, pluck, "umap_model", "embedding"),
        embed = map(embed, data.frame),
        clusterdat = map(data, pluck, "probs"),
        clusterdat = map2(clusterdat, embed, bind_cols),
        clusterdat = map(clusterdat, pivot_longer, -c(eid, X1, X2), names_to = "cluster", values_to = "prob"),
        clusterdat = map(clusterdat, group_by, eid),
        clusterdat = map(clusterdat, slice_max, prob, with_ties = FALSE)
    ) %>%
    select(cohort, sex, clusterdat) %>%
    unnest(clusterdat)
head(closest_cluster)
```

```{r}
options(repr.plot.width = 10, repr.plot.height = 2)
closest_cluster %>%
    mutate(
        cohort = recode_factor(
            factor(cohort), 
            "ukb" = "UKBiobank",
            "ghs" = "GHS",
            "rotterdam" = "Rotterdam",
            "maastricht" = "Maastricht"
        )
    ) %>%
    ggplot(aes(X1, X2)) +
    geom_point(aes(fill = cluster), show.legend = FALSE,
               size = .2, shape = 21, color = alpha("magenta", 0.2), stroke = .01, alpha = .8) +
    ggh4x::facet_nested(~cohort + sex, scales = "free", independent = "all") +
    theme(
        plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black"),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = "black", color = "white"),
        strip.text = element_text(color = "white"),
        axis.text = element_blank(),
        axis.ticks = element_blank()
    ) +
    labs(x = "UMAP1", y = "UMAP2")
```

## Discovery data

```{r}
load("../data/ukb/strat_dat.RData")
```

```{r}
map(strat_dat, head)
```

## Residualization of discovery data by each model

```{r}
BIOMARKERS <- c("whr", "sbp", "dbp", "alt", "scr", "crp", "hdl", "tg", "ldl", "fg")
```

```{r}
COVARIATES <- c("bmi", "age", "smoking")
```

```{r}
predicttab <- regcoefs %>%
    nest(coefdat = -c(cohort, sex, Biomarker)) %>%
    mutate(
        coefmat = map(coefdat, ~matrix(.x$Estimate, ncol = 1, dimnames = list(.x$term, "Estimate"))),
        coefmat = map(coefmat, ~.x[c("(Intercept)", COVARIATES), "Estimate", drop = FALSE]),
        observedval = map2(sex, Biomarker, ~pull(strat_dat[[.x]], .y)),
        predictormat = map(sex, ~as.matrix(cbind(Intercept = 1, strat_dat[[.x]][,COVARIATES]))),
        predictedval = map2(predictormat, coefmat, ~as.vector(.x %*% .y)),
        residval = map2(observedval, predictedval, ~.x - .y),
        MeanRes = map_dbl(residval, mean),
        SDRes = map_dbl(residval, sd)
    )
print(predicttab)
```

```{r}
residtab <- predicttab %>%
    mutate(scaleresidval = pmap(list(residval, MeanRes, SDRes), ~(..1 - ..2) / ..3)) %>%
    select(cohort, sex, Biomarker, scaleresidval) %>%
    pivot_wider(names_from = Biomarker, values_from = scaleresidval) %>%
    mutate(eid = map(sex, ~strat_dat[[.x]]$eid)) %>%
    unnest(-c(cohort, sex)) %>%
    select(cohort, sex, eid, all_of(BIOMARKERS))
head(residtab)
```

## Cluster validation

```{r}
cluster_results <- result_tab %>%
    transmute(
        cohort,
        clusdat = map(data, "Cluster_results"),
        sex = map(clusdat, names)
    ) %>%
    unnest(c(sex, clusdat)) %>%
    mutate(clusdat = map(clusdat, "clusters"))
print(cluster_results)
```

```{r}
clusprobdat <- cluster_results %>%
    mutate(
        residdat = map2(cohort, sex, ~filter(residtab, cohort == .x, sex == .y)),
        probdat = map2(
            residdat, clusdat,
            function(DAT, CLUSDAT){
                X <- DAT[,BIOMARKERS]
                mus <- map(CLUSDAT, "center")
                mus <- map(mus, function(mu){ mu[BIOMARKERS] })
                covmats <- map(CLUSDAT, "cov")
                covmats <- map(covmats, function(covmat){ covmat[BIOMARKERS, BIOMARKERS] })
                wts <- map_dbl(CLUSDAT, "weight")
                pdfs <- mapply(function(mu, covmat){ mvtnorm::dmvnorm(X, mu, covmat) }, mus, covmats)
                L <- pdfs %*% diag(wts)
                colnames(L) <- colnames(pdfs)
                data.frame(eid = DAT$eid, L / rowSums(L))
            }
        )
    )
print(clusprobdat)
```

```{r}
validationtab <- clusprobdat %>%
    transmute(
        cohort, sex, 
        probdat = map(probdat, pivot_longer, -eid, names_to = "cluster", values_to = "prob")
    ) %>%
    {
        ## Discovery vs. validation cohorts
        refdat <- filter(., cohort == "ukb")
        testdat <- filter(., cohort != "ukb")
        inner_join(refdat, testdat, by = "sex", suffix = c("_ref", "_test"))
    } %>%
    mutate(
        ## Individuals with high certainty of cluster allocation in discovery model
        probdat_ref = map(probdat_ref, filter, prob >= .8),
        probdat = map2(probdat_ref, probdat_test, inner_join, by = "eid", suffix = c("_ref", "_test"))
    ) %>%
    select(sex, cohort_ref, cohort_test, probdat) %>%
    unnest(probdat) %>%
    filter(cluster_ref != "cluster_0")
head(validationtab)
```

```{r}
medianprobtab <- validationtab %>%
    group_by(sex, cohort_ref, cohort_test, cluster_ref, cluster_test) %>%
    summarise(MedProb = median(prob_test), .groups = "drop")
head(medianprobtab)
```

```{r}
validatedtab <- medianprobtab %>%
    filter(MedProb >= .8)
head(validatedtab)
```

```{r}
options(repr.plot.width = 7, repr.plot.height = 4)
validatedtab %>%
    transmute(
        sex,
        from = gsub("cluster_", "U:", cluster_ref),
        to = case_match(
            cohort_test,
            "maastricht" ~ gsub("cluster_", "M:", cluster_test),
            "rotterdam" ~ gsub("cluster_", "R:", cluster_test),
            "ghs" ~ gsub("cluster_", "G:", cluster_test)
        )
    ) %>%
    group_by(sex) %>%
    nest %>%
    mutate(
        net = map(data, ~igraph::graph_from_data_frame(.x, directed = FALSE)),
        netplot = map2(
            net, sex,
            ~.x %>%
                ggraph::ggraph(layout = "igraph", algorithm = "nicely") +
                ggraph::geom_edge_link() +
                ggraph::geom_node_point(aes(fill = gsub("^([A-Z]):.+$", "\\1", name)), 
                                        shape = 21, size = 7) +
                scale_fill_viridis_d(
                    option = "A", begin = .25, end = .75, direction = -1,
                    labels = c(U = "UKB", M = "TMS", R = "RS", G = "GHS"),
                    guide = guide_legend(override.aes = list(size = 2))
                ) +
                ggraph::geom_node_text(aes(label = name), color = "white", fontface = "bold", size = 2) +
                labs(title = .y, fill = "Cohort")
        )
    ) %>%
    pull(netplot) %>%
    patchwork::wrap_plots(nrow = 1, guides = "collect") &
    theme(legend.position = "bottom")
```

### Cluster mapper

```{r}
cluster_mapper <- validatedtab %>%
    select(sex, cluster_ref, cohort_test, cluster_test) %>%
    group_by(sex, cluster_ref) %>%
    filter(n() == 3) %>%
    mutate(
        validclus_id = paste0("VC", cur_group_id())
    ) %>%
    ungroup %>%
    arrange(sex, validclus_id)
head(cluster_mapper)
```

```{r}
cluster_mapper <- cluster_mapper %>%
    {
        refdat <- transmute(., sex, validclus_id, cohort = "ukb", cluster_id = cluster_ref)
        refdat <- unique(refdat)
        testdat <- transmute(., sex, validclus_id, cohort = cohort_test, cluster_id = cluster_test)
        bind_rows(refdat, testdat)
    } %>%
    arrange(sex, validclus_id)
head(cluster_mapper)
```

```{r}
options(repr.plot.width = 12, repr.plot.height = 3)
cluster_mapper %>%
    inner_join(cluster_results) %>%
    transmute(
        cohort, sex, validclus_id, cluster_id,
        centers = map2(clusdat, cluster_id, ~tibble(trait = BIOMARKERS, center = .x[[.y]][["center"]][BIOMARKERS]))
    ) %>%
    unnest(centers) %>%
    ggplot(aes(center, trait)) +
    geom_vline(xintercept = 0, lty = "dashed") +
    geom_linerange(aes(xmin = 0, xmax = center, group = cohort, color = cohort),
                 position = position_dodge(width = .5)) +
    geom_point(aes(group = cohort, color = cohort),
               position = position_dodge(width = .5)) +
    ggh4x::facet_nested(~ sex + validclus_id) +
    theme_bw() +
    theme(legend.position = "top")
```

## Naming clusters based on centers

```{r}
cluster_namer <- cluster_mapper %>%
    mutate(
        validclus_name = case_match(
            validclus_id,
            "VC1" ~ "DHT",
            "VC2" ~ "DHG",
            "VC3" ~ "DIS",
            "VC4" ~ "DAL",
            "VC5" ~ "DLT",
            "VC6" ~ "DAL",
            "VC7" ~ "DLT",
            "VC8" ~ "DHG",
            "VC9" ~ "DIS"
        )
    )
head(cluster_namer)
```

```{r}
write_tsv(cluster_namer, "../data/cluster_namer.tsv")
```

```{r}
options(repr.plot.width = 7, repr.plot.height = 5)
cluster_mapper %>%
    inner_join(cluster_results) %>%
    transmute(
        sex, cohort, validclus_id, cluster_id,
        centers = map2(clusdat, cluster_id, ~tibble(trait = BIOMARKERS, center = .x[[.y]][["center"]][BIOMARKERS]))
    ) %>%
    unnest(centers) %>%
    inner_join(cluster_namer) %>%
    mutate(
        cohort = stringr::str_replace_all(
            cohort,
            c(
                "ukb" = "UKB",
                "maastricht" = "MS",
                "rotterdam" = "RS",
                "ghs" = "GHS"
            )
        )
    ) %>%
    split(f = .$sex) %>%
    map(
        ~ggplot(.x, aes(center, trait)) +
            geom_vline(xintercept = 0, lty = "dashed") +
            geom_linerange(aes(xmin = 0, xmax = center, group = cohort, color = cohort),
                           position = position_dodge(width = .5)) +
            geom_point(aes(group = cohort, color = cohort),
                       position = position_dodge(width = .5)) +
            ggh4x::facet_nested(~ sex + validclus_name) +
            theme_bw()
    ) %>%
    patchwork::wrap_plots(ncol = 1, guides = "collect") &
    theme(legend.position = "top", legend.title = element_blank(), axis.title = element_blank())
```

## Saving validated clustering model

```{r}
valid_cluspars <- cluster_results %>%
    filter(cohort == "ukb") %>%
    inner_join(cluster_namer) %>%
    transmute(
        sex, validclus_name,
        validclus_centers = unname(map2(clusdat, cluster_id, ~.x[[.y]][["center"]])),
        validclus_covmats = unname(map2(clusdat, cluster_id, ~.x[[.y]][["cov"]])),
        validclus_weights = map2_dbl(clusdat, cluster_id, ~.x[[.y]][["weight"]])
    )
print(valid_cluspars)
```

Adding concordant cluster:

```{r}
concordantclus <- valid_cluspars %>%
    group_by(sex) %>%
    summarise(
        validclus_name = "BC",
        validclus_centers = list(setNames(rep(0, length(BIOMARKERS)), BIOMARKERS)),
        validclus_covmats = list(structure(diag(rep(1, length(BIOMARKERS))), 
                                           dimnames = list(BIOMARKERS, BIOMARKERS))),
        validclus_weights = 1 - sum(validclus_weights)
    )
print(concordantclus)
```

```{r}
valid_cluspars <- valid_cluspars %>%
    bind_rows(concordantclus) %>%
    arrange(sex, desc(validclus_weights))
print(valid_cluspars)
```

```{r}
valid_cluspars <- valid_cluspars %>%
    nest(clusmod = -sex)
print(valid_cluspars)
```

Regression model coefficients:

```{r}
RGC <- regcoefs %>% 
    filter(cohort == "ukb") %>% 
    transmute(sex, Biomarker, term = gsub("\\(|\\)", "", term), Estimate) %>%
    pivot_wider(names_from = term, values_from = Estimate)
head(RGC)
```

Standard deviation of residuals:

```{r}
RESIDSD <- predicttab %>% 
    filter(cohort == "ukb") %>% 
    select(sex, Biomarker, SDRes)
head(RESIDSD)
```

```{r}
LMMODELDF <- inner_join(RGC, RESIDSD)
head(LMMODELDF)
```

```{r}
LMMODELDF <- LMMODELDF %>%
    nest(residmod = -sex)
print(LMMODELDF)
```

Final per-sex model

```{r}
validclusmod <- inner_join(LMMODELDF, valid_cluspars)
print(validclusmod)
```

```{r}
save(validclusmod, file = "../data/validclusmod.RData")
```