### 2025-11-12 scripts/03_CentBased.Rmd

---
title: An R Markdown document converted from "03_CentBased.ipynb"
output: html_document
---

# Centroid-based clustering

```{r}
library(readr)
library(dplyr, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
library(purrr)
library(ggplot2)
```

```{r}
options(repr.plot.res = 300)
```

## Loading data

```{r}
load("../data/validclusmod.RData")
```

```{r}
print(validclusmod)
```

```{r}
validclusmod <- validclusmod %>%
    mutate(
        umapmod = map(
            sex,
            ~uwot::load_uwot(paste0("../data/ukb/umapmod_", .x))
        )
    )
print(validclusmod)
```

```{r}
validclusmod <- validclusmod %>%
    mutate(
        residtab = map(
            umapmod,
            ~map(
                1:.x$nn_index$ann$getNItems(),
                function(i){
                    .x$nn_index$ann$getItemsVector(i-1)
                }
            )
        ),
        residtab = map(residtab, do.call, what = rbind)
    )
print(validclusmod)
```

## Centroid based clustering

```{r}
gausmixres <- validclusmod %>%
    transmute(
        sex,
        choosemod = map(residtab, mclust::mclustBIC, G = 1:20, modelNames = "VVV")
    )
```

```{r}
print(gausmixres)
```

```{r}
options(repr.plot.height = 3, repr.plot.width = 7)
gausmixres %>%
    mutate(
        choosemod = map(
            choosemod,
            ~data.frame(
                Nclus = 1:20,
                BIC = .x[,1]
            )
        )
    ) %>%
    unnest(choosemod) %>%
    ggplot(aes(Nclus, BIC)) +
    geom_line() +
    geom_point() +
    facet_wrap(~sex, scales = "free") +
    theme_bw()
```

```{r}
gmbest <- validclusmod %>%
    transmute(
        sex,
        mod = map(
            residtab, 
            ~{
                mclustBIC <- mclust::mclustBIC
                mclust::Mclust(.x, G = 6, modelNames = "VVV")
            }
        )
    )
```

```{r}
options(repr.plot.height = 4, repr.plot.width = 8)
gmbestcent <- gmbest %>%
    transmute(
        sex,
        meandat = map(mod, ~.x$parameters$mean)
    )
print(gmbestcent)
```

```{r}
options(repr.plot.height = 4, repr.plot.width = 8)
gmbestcent %>%
    mutate(
        meandat = map(
            meandat,
            ~{
                res <- t(.x)
                colnames(res) <- c("whr", "sbp", "dbp", "alt", "scr", "crp", "hdl", "tg", "ldl", "fg")
                data.frame(Cluster = paste0("Cluster", 1:nrow(res)), res)
            }
        )
    ) %>%
    unnest(meandat) %>%
    pivot_longer(-c(sex, Cluster), names_to = "trait", values_to = "residval") %>%
    mutate(trait = toupper(trait)) %>%
    ggplot(aes(residval, trait)) +
    geom_segment(aes(xend = 0, yend = trait)) +
    geom_vline(xintercept = 0, lty = "dashed") +
    geom_point() +
    facet_grid(sex ~ Cluster) +
    labs(x = "Residuals - SD units", y = NULL) +
    theme_bw()
```

```{r}
options(repr.plot.height = 2.5, repr.plot.width = 4.5)
gmbest %>%
    transmute(
        sex,
        weightdat = map(
            mod,
            ~data.frame(Cluster = paste0("Cluster", 1:.x$G), 
                        Weight = .x$parameters$pro)
        )
    ) %>%
    unnest(weightdat) %>%
    ggplot(aes(Weight, reorder(Cluster, desc(Cluster)))) +
    geom_col() +
    facet_wrap(~sex) +
    scale_x_continuous(labels = \(x) 100 * x) +
    theme_bw() +
    labs(x = "Weight (%)", y = NULL)
```

```{r}
gmbest %>%
    transmute(
        sex,
        RelEnt = map_dbl(
            mod,
            ~{
                pp <- pmax(as.matrix(.x$z), .Machine$double.xmin)
                ent <- -sum(rowSums(pp * log(pp)))
                1 - ent / (nrow(pp) * log(ncol(pp)))
            }
        )
    )
```

```{r}
embeddf <- validclusmod %>%
    transmute(sex, umapmod = map(umapmod, ~data.frame(.x$embedding))) %>%
    unnest(umapmod)
head(embeddf)
```

```{r}
centembed <- validclusmod %>%
    inner_join(gmbestcent, by = "sex") %>%
    transmute(
        sex,
        meandat = map2(
            meandat, umapmod,
            ~data.frame(uwot::umap_transform(t(.x), .y))
        )
    ) %>%
    unnest(meandat)
head(centembed)
```

```{r}
options(repr.plot.width = 5, repr.plot.height = 3)
embeddf %>%
    ggplot(aes(X1, X2)) +
    geom_point(shape = ".", alpha = .1) +
    geom_point(data = centembed, color = "red") +
    facet_wrap(~sex) +
    theme_bw() +
    labs(x = "UMAP1", y = "UMAP2")
```